\chapter{EFT parametrisation: MC generator details}\label{app:generator_options}
\vspace{-.5cm}
This Appendix describes the technical details in deriving the EFT parametrisations, namely the options used to configure the MC generators. Firstly, the treatment of $m_{\rm{Z}}$ in the HEL interpretation is described. Following this, a crib sheet is provided in Table~\ref{tab:generator_options}, which lists all the relevant details.
\vspace{-.3cm}
\section{The treatment of $m_{\rm{Z}}$ in the HEL interpretation}
The nominal HEL model is uniquely defined by the input parameters: $m_{\rm{W}}$, $\alpha_{\rm{EM}}$ and $G_F$, where $m_{\rm{W}}$ is the W boson mass, $\alpha_{\rm{EM}}$ is the electromagnetic coupling constant, and $G_F$ is the Fermi constant. In this input parameter scheme, the Z boson mass, $m_{\rm{Z}}$, is dependent on the HEL parameters, $c_T$, $c_{WW}$, $c_B$ and $c_A$, according to equation \ref{eq:hel_mz},
\begin{equation}\label{eq:hel_mz}
    m_{\rm{Z}}^2(\vec{c}) = m_{Z,\rm{SM}}^2 \Big[ 1-c_T+\frac{8\,c_A\,\sin^4(\theta_W)+2\,c_{WW}\,\cos^2(\theta_W)+c_B\,\sin^2(\theta_W)}{\cos^2(\theta_W)} \Big],
\end{equation}

\noindent
where $m_{Z,\rm{SM}}$ is the Z boson mass in the SM, and $\theta_W$ is the Weinberg angle. In the interpretation documented in this thesis, the HEL model has been adapted to remove the $\vec{c}$ dependence of $m_{\rm{Z}}$, and instead fix its value to the SM prediction. This reflects the fact that, although there is no explicit Z boson mass measurement entering the combination, it is well measured experimentally and therefore it is not physical to consider large variations in its value. In fact, the chosen operator subset, $\{\mathcal{O}\}$, does not include $\mathcal{O}_T$, and $\mathcal{O}_{WW}$ and $\mathcal{O}_B$ are fit together in the combination of parameters, $c_{WW}-c_B$, which equation \ref{eq:hel_mz} does not explicitly depend on. The operator $\mathcal{O}_A$ is included in the subset, however the $c_A$ dependence is small ($\propto \sin^4(\theta_W)$), such that over the allowed range in $c_A$ there is a negligible shift in the $m_{\rm{Z}}$ value, well within its measured uncertainty. All in all, this means the treatment of $m_{\rm{Z}}$ in the interpretation has a negligible effect.

\begin{table}[htb!]
  \centering
  \footnotesize
  \renewcommand{\arraystretch}{1.5}
  \setlength{\tabcolsep}{6pt}
  \caption[MC generator details for deriving the EFT parametrisation]
  {
    Crib sheet for the MC generator details used when deriving the EFT parametrisation.
  }
  \label{tab:generator_options}
%   \hspace*{-1cm}
  \input{Tables/eft/generator_options}
%   \hspace*{-1cm}
\end{table}